How fluid-mechanical erosion creates anisotropic porous media

### 2022-06-29 docs/PhysicsPaper/StokesErosion.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quadrature for Cauchy integrals}
The accuracy of our method is determined by the accuracy of the
quadrature rule applied to equation~\eqref{eqn:FredBIE}. Since we have
written the Stokes double-layer potential
velocity~\eqref{eqn:velocityCauchy}, deformation
tensor~\eqref{eqn:deformationCauchy}, and
vorticity~\eqref{eqn:vorticityCauchy} as a sum of Cauchy integrals and
their derivatives, the accuracy of the method depends on the accuracy of
computing a general Cauchy integral
\begin{align}
  v(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x-y} \, dy.
  \label{eqn:cauchy1}
\end{align}
Here we describe a quadrature formulae that was first used to
approximate analytic functions~\cite{ioa-pap-per1991}, and then extended
to Stokes layer potentials~\cite{bar-wu-vee2015}. The quadrature method
requires the boundary data of the analytic function~\eqref{eqn:cauchy1}
which satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  v^{-}(x_0) = \lim_{\substack{x \rightarrow x_0 \\ x \in \Omega}} \int_{\gamma}
    \frac{\eta(y)}{x-y}\, dy = -\frac{1}{2}\eta(x_0) - 
    \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x_0-y} \, dy,
    \quad x_0 \in \gamma,
    \label{eqn:SP}
\end{align}
where the last integral is interpreted in the principal value sense.
Here we are assuming that $\Omega$ is the bounded region interior to
$\gamma$. Once $v^{-}$ is calculated, $v(x)$ and its derivatives can be
determined by its boundary data alone using the Cauchy Integral Theorem
\begin{subequations}
  \label{eqn:cauchy}
  \begin{alignat}{3}
  \label{eqn:cauchyv}
  v(x) &= \frac{1}{2\pi i}\int_{\gamma} 
    \frac{v^{-}(y)}{y-x} \,dy, \\
  v'(x) &= \frac{1}{2\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^2} \, dy, \\
  v''(x) &= \frac{1}{\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^3} \, dy.
  \end{alignat}
\end{subequations}

@@

% SNIPPET
%This has the advantage that only the one-dimensional boundary of the domain must be discretized, and, with appropriate quadrature formulas and fast summation methods, the result is a high-fidelity numerical simulation with near-optimal computational complexity.

@@

With the flow and shear stresses computed, we seek to evolve the boundaries of erodable bodies. For numerical stability, we modify~\eqref{Vn0} to include a smoothing term that depends on local curvature $\kappa$, as in previous work~\cite{quaife2018boundary}
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\pi} \kappa - 1 \right).
\end{align}
The last term is a smoothing term that has strength $\epsilon \ll 1$ and scales with spatial average of the shear stress $\langle\abs{\tau}\rangle$. $L$ indicates the total arc length of the body, and, inside the parenthesis, the mean curvature is subtracted so that this term preserves area. As such, the only source of material loss is the first term, $\CE \, \abs{\tau}$, which represents the shear-dependent erosion law. In addition, we apply a narrow Gaussian filter to the distribution $\abs{\tau}$ to further improve stability.

@@

% COPIED STRAIGHT FROM JFM 2020
%We use the time stepping method outlined in our previous work~\citep[see][section 3.3]{qua-moo2018} which we briefly summarize here.  The erosion rate loses differentiability if the shear stress changes sign, and this leads to corners developing on $\gamma$ and numerical instabilities.  Therefore, we modify the erosion rate, $V_\nn$ in equation~\eqref{eqn:erosionModel}, with
%\begin{align}
% \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(\frac{L}{2\pi} \kappa - 1 \right),
%\end{align}
%where $\epsilon \ll 1,$ $\langle \cdot \rangle$ is the spatial average, $L$ is the length of $\gamma$, and $\kappa$ is the curvature of $\gamma$.  The new erosion model penalizes regions of high curvature, but does not change the total length of each body.  Moreover, to increase the overall stability of the method, a narrow Gaussian filter is applied to the erosion rate at each time step.

%Rather than tracking the $(x,y)$ coordinates, the {\thL} coordinates are tracked. In addition, tangential velocity fields are used to maintain an equispaced discretization. Time stepping is performed with a second-order Implicit-Explicit Runge-Kutta method. In particular, the diffusive term corresponding to the curvature penalization term is discretized implicitly, and all other terms, which are non-stiff, are treated explicitly.  By using this time stepping method in conjunction with the Barycentric rule, we stably evolve the eroding bodies.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quadrature for Cauchy integrals}
The accuracy of our method is determined by the accuracy of the
quadrature rule applied to equation~\eqref{eqn:FredBIE}. Since we have
written the Stokes double-layer potential
velocity~\eqref{eqn:velocityCauchy}, deformation
tensor~\eqref{eqn:deformationCauchy}, and
vorticity~\eqref{eqn:vorticityCauchy} as a sum of Cauchy integrals and
their derivatives, the accuracy of the method depends on the accuracy of
computing a general Cauchy integral
\begin{align}
  v(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x-y} \, dy.
  \label{eqn:cauchy1}
\end{align}
Here we describe a quadrature formulae that was first used to
approximate analytic functions~\cite{ioa-pap-per1991}, and then extended
to Stokes layer potentials~\cite{bar-wu-vee2015}. The quadrature method
requires the boundary data of the analytic function~\eqref{eqn:cauchy1}
which satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  v^{-}(x_0) = \lim_{\substack{x \rightarrow x_0 \\ x \in \Omega}} \int_{\gamma}
    \frac{\eta(y)}{x-y}\, dy = -\frac{1}{2}\eta(x_0) - 
    \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x_0-y} \, dy,
    \quad x_0 \in \gamma,
    \label{eqn:SP}
\end{align}
where the last integral is interpreted in the principal-value sense.
Here we are assuming that $\Omega$ is the bounded region interior to
$\gamma$. Once $v^{-}$ is calculated, $v(x)$ and its derivatives can be
determined by its boundary data alone using the Cauchy Integral Theorem
\begin{subequations}
  \label{eqn:cauchy}
  \begin{alignat}{3}
  \label{eqn:cauchyv}
  v(x) &= \frac{1}{2\pi i}\int_{\gamma} 
    \frac{v^{-}(y)}{y-x} \,dy, \\
  v'(x) &= \frac{1}{2\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^2} \, dy, \\
  v''(x) &= \frac{1}{\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^3} \, dy.
  \end{alignat}
\end{subequations}

@@

@@

With the flow computed, we extract the vorticity which reduces to shear stress, $\tau$, on solid boundaries. Next, we seek to evolve the boundaries of these erodable bodies. For numerical stability, we modify erosion law \eqref{Vn0} to include a smoothing term that depends on local curvature $\kappa$~\cite{quaife2018boundary}
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\pi} \kappa - 1 \right).
\end{align}
The last term is a smoothing term that has strength $\epsilon \ll 1$ and scales with the spatial average of the shear stress $\langle\abs{\tau}\rangle$. $L$ indicates the total arc length of the body, and, inside the parenthesis, the mean curvature is subtracted so that this term preserves area. As such, the only source of material loss is the first term, $\CE \, \abs{\tau}$, representing the shear-dependent erosion law \eqref{Vn0}. In addition, we apply a narrow Gaussian filter to the distribution $\abs{\tau}$ to further improve stability.

@@
